Photo-D with LSST: stellar photometric distances out to the edge of the Galaxy

### 2024-08-29 paper/pipeline.tex
minor fixed in Sec. 4

The scale of the dataset expected from LSST -- O(10Bn) objects -- makes this
computation challenging to perform on a single core or computer. Ideally,
one would wish to execute the computation in parallel on many (hundreds) of cores,
likely housed on a number of separate nodes. This requires a distributed
processing framework capable of orchestrating such a solution.

@@

For our work, we've adopted the \href{https://lsdb.readthedocs.io/}{\em Large Survey Database} (LSDB; \cite{2024AAS...24326109W})
framework. LSDB is a distributed computing framework written in Python,
which enables multi-node and multi-core computation on large (PB-scale),
partitioned, catalog datasets.  It extends the concepts of the LSD toolkit
\citep{2011AAS...21743319J}
to astronomical catalogs where the distribution of objects on the sky is
very uneven -- such as is the case for Galactic star counts.  Leveraging
broadly adopted community libraries such as astropy, Pandas, and Dask, LSDB
presents a user-friendly API on which it is possible to build our pipeline.

@@

This workflow is particularly efficient and well-suited for the scientific objectives addressed by {\em PhotoD}, as it leverages the variable resolution of the map based on the star density within each region. For instance, in the galactic plane where star density is higher, smaller pixels provide higher resolution, while in the halo, where star density is lower, larger pixels suffice.

@@

We have performed several benchmarks to estimate the performance of the PhotoD code. In our initial tests based on the SDSS Stripe 82 data we found that the distances for a billion stars can be estimated in about 100 days on a Mac M1 Pro laptop using one core (10 milliseconds per star).

In further testing, we used outputs from the TRILEGAL simulation and found that the parallel processing performed with LSDB scaled well with respect to the single core values. On a local cluster node with dual AMD EPYC 9474F 48-core processors we were able to achieve per-star figures similar to single-core processing, therefore reducing the required processing time to about a week. Our ongoing tests are showing that when all 10 nodes (400 cores) of the Ruđer Bošković Institute's Narval cluster are used, we may be able to estimate distances for about 10 billion stars in about two days.

>>>>

The scale of the dataset expected from LSST -- O(10Bn) objects -- makes this
computation challenging to perform on a single core or computer. Ideally,
one would wish to execute the computation in parallel on many (hundreds of) cores,
likely housed on a number of separate nodes. This requires a distributed
processing framework capable of orchestrating such a solution.

@@

For our work, we have adopted the \href{https://lsdb.readthedocs.io/}{\em Large Survey Database} 
\citep[LSDB;][]{2024AAS...24326109W} framework. LSDB is a distributed computing framework written in Python,
which enables multi-node and multi-core computation on large (PB-scale),
partitioned, catalog datasets.  It extends the concepts of the LSD toolkit
\citep{2011AAS...21743319J}
to astronomical catalogs where the distribution of objects on the sky is
very uneven -- such as is the case for Galactic star counts.  Leveraging
broadly adopted community libraries such as astropy, Pandas, and Dask, LSDB
presents a user-friendly API on which it is possible to build our pipeline.

@@

This workflow is particularly efficient and well-suited for the scientific objectives addressed by {\em PhotoD}, as it leverages the variable resolution of the map based on the star density within each region. For instance, in the Galactic plane where star density is higher, smaller pixels provide higher resolution, while in the halo, where star density is lower, larger pixels suffice.

@@

We have performed several benchmarks to estimate the performance of the PhotoD code. In our initial tests based on the SDSS Stripe 82 data we found that the distances for a billion stars can be estimated in about 100 days on a Mac M1 Pro laptop using one core (corresponding to 10 milliseconds per star).

In further testing, we used outputs from the TRILEGAL simulation and found that the parallel processing performed with LSDB scaled well with respect to the single core values. On a local cluster node with dual AMD EPYC 9474F 48-core processors, we were able to achieve per-star figures similar to single-core processing, therefore reducing the required processing time to about a week. Our ongoing tests are showing that when all 10 nodes (400 cores) of the Ruđer Bošković Institute's Narval cluster are used, we may be able to estimate distances for 10 billion stars in about two days.
